---
title: "1-22"
author: "Andrew Wu"
date: "2025-01-30"
output: html_document
---

Suppose we flip a coin $n$ times and let $p$ denote the probability of heads. Let $X$ be the number of heads. We call $X$ a binomial random variable. Intuition suggests $X$ will be close to $np$. We can repeat this experiment many times and average the $X$ values. Carry out a simulation and compare the average of the $X$'s to $np$. Try this for $p = .3$ and $n = 10, n = 100,$ and $n = 1000$.

```{r}
set.seed(130)
```

```{r}
n <- 10 # number of coinflips
p <- 0.3 # probability of heads
m <- 1000 # number of times we repeat the experiment
```

```{r}
Xs <- numeric(m) # create an m-dimensional vector
for (a in 1:m) {
  X <- sum(rbinom(n,1,p)) # generate a binomial RV
  Xs[a] <- X # throw it into Xs
}
```

```{r}
mean(Xs)
```
Here $np = 10\cdot 0.3 = 3$ and the average of the X's is 2.975.

```{r}
n <- 100
for (a in 1:m) {
  X <- sum(rbinom(n,1,p)) # generate a binomial RV
  Xs[a] <- X # throw it into Xs
}
mean(Xs)
```

Here $np = 30$ and the average of the X's is 30.052.

```{r}
n <- 1000
for (a in 1:m) {
  X <- sum(rbinom(n,1,p)) # generate a binomial RV
  Xs[a] <- X # throw it into Xs
}
mean(Xs)
```
Here $np = 300$ and the average of the X's is 300.045.
