---
title: "1-23"
author: "Andrew Wu"
date: "2025-01-31"
output: html_document
---

Consider tossing a fair die. Let $A = \{2,4,6\}$ and $B = \{1,2,3,4\}$. Then $\mathbb{P}(A) = 1/2, \mathbb{P}(B) = 2/3,$ and $\mathbb{P}(AB) = 1/3$. Evidently events $A$ and $B$ are independent. Simulate draws from the sample space and verify that $\hat{\mathbb{P}}(AB) = \hat{\mathbb{P}}(A) \hat{\mathbb{P}}(B)$, where $\hat{\mathbb{P}}(A)$ is the proportion of times $A$ occurred in the simulation (and similar for $\hat{\mathbb{P}}(AB)$ and $\hat{\mathbb{P}}(B)$.)

Next, find two events $A$ and $B$ are not independent, and compute the same values. Compare the calculated values to their theoretical values. Report your results and interpret.

```{r}
set.seed(127)

# create sets A and B
A <- array(c(2,4,6)) 
B <- array(c(1,2,3,4))

# set total rolls, initialize success variables
n <- 1000
totalAs <- 0
totalBs <- 0
totalABs <- 0

# simulate rolls from sample space
rolls <- sample(1:6,n, replace='True')
for (r in rolls) {
  # check whether each roll is in A, B, or AB
  if (r %in% A) {
    totalAs <- totalAs + 1
  }
  if (r %in% B) {
    totalBs <- totalBs + 1
  }
  if ((r %in% A) && (r %in% B)) {
    totalABs <- totalABs + 1
  }
}
```

```{r}
probAhat <- totalAs/n
probBhat <- totalBs/n
probABhat <- totalABs/n

print(paste("The proportion of times A was observed is", probAhat))
print(paste("The proportion of times B was observed is", probBhat))
print(paste("The proportion of times AB was observed is", probABhat))
print(paste("The product of 0.483 and 0.676 is 0.326, which is close to 0.322."))

```

To check for non-independent events, simply take $A = \{1,2,3\}$ and $B = \{4,5,6\}$. 

```{r}
set.seed(127)

# create sets A and B
A <- array(c(1,2,3)) 
B <- array(c(4,5,6))

# initialize success variables
totalAs <- 0
totalBs <- 0
totalABs <- 0

# simulate rolls from sample space
rolls <- sample(1:6,n, replace='True')
for (r in rolls) {
  # check whether each roll is in A, B, or AB
  if (r %in% A) {
    totalAs <- totalAs + 1
  }
  if (r %in% B) {
    totalBs <- totalBs + 1
  }
  if ((r %in% A) && (r %in% B)) {
    totalABs <- totalABs + 1
  }
}
```

```{r}
probAhat <- totalAs/n
probBhat <- totalBs/n
probABhat <- totalABs/n

print(paste("The proportion of times A was observed is", probAhat))
print(paste("The proportion of times B was observed is", probBhat))
print(paste("The proportion of times AB was observed is", probABhat))
print(paste("The product of 0.519 and 0.481 is not close to 0."))
```

