---
title: '11.4'
author: "Andrew Wu"
date: "2025-06-29"
output: html_document
---

Suppose that $50$ people are given a placebo and $50$ are given a new treatment. $30$ placebo patients show improvement while $40$ treated patients show improvement. Let $\tau = p_2 - p_1$, where $p_2$ is the probability of improving under treatment and $p_1$ is the probability of improving under placebo.

(b) Find the standard error and $90$ percent confidence interval using the parametric bootstrap.
(c) With the prior $f(p_1,p_2) = 1$, use simulation to find the posterior mean and posterior $90$ percent interval for $\tau$.
(e) Use simulation to find the posterior mean and posterior $90$ percent interval for $\psi$.

```{r}
# noting given information
n <- 50
s1 <- 30
s2 <- 40
```

We found in the previous part the MLEs of $p_1, p_2,$ and $\tau$.

```{r}
p_1_hat <- s1/n
p_2_hat <- s2/n
tau_hat <- p_2_hat - p_1_hat
```

Now we bootstrap.

```{r}
B <- 10000
tau_bootstrapped <- c()
for (i in 1:B) {
  # X_i := # of ppl who show improvement under placebo
  X_i <- sum(rbinom(n,1,p_1_hat))
  # Y_i := ... under new treatment
  Y_i <- sum(rbinom(n,1,p_2_hat))
  tau_bootstrapped[i] <- Y_i/n - X_i/n
}
```

Now we calculate the estimated standard error and a 90% confidence interval.

```{r}
se_boot <- sd(tau_bootstrapped)
se_boot
```

```{r}
CI_boot <- quantile(tau_bootstrapped, probs = c(0.05, 0.95))
CI_boot
```

With posterior $f(p_1,p_2) = 1$, we have by Bayes' theorem that $$f(p_1,p_2 | x_i,y_i) \propto f(x_i,y_i | p_1, p_2) f(p_1,p_2).$$

Thus $$f(p_1,p_2 | x_i,y_i) \propto p_1^{x_i}(1-p_1)^{n - x_i} p_2^{y_i}(1-p_2)^{n - y_i}.$$

Note also that $$f(p_1,p_2 | x_i,y_i) = f(p_1 | x_i) f(p_2 | y_i)$$ so $p_1, p_2$ are independent under the posterior. Moreover, $p_1 | x_i \sim \text{Beta} (x_i+1, n - x_i + 1)$ and $p_2 | y_i \sim \text{Beta} (y_i + 1, n - y_i + 1)$.

That means we can simulate as follows:

```{r}
# number of simulations of tau
N <- 1000

# simulate p1s and p2s, to get simulated taus
p1_sim <- rbeta(N, s1 + 1, n - s1 + 1)
p2_sim <- rbeta(N, s2 + 1, n - s2 + 1)
tau_sim <- p2_sim - p1_sim

# obtain posterior mean
tau_post_mean <- mean(tau_sim)
tau_post_mean

# obtain CI
tau_post_CI <- quantile(tau_sim, probs = c(0.05, 0.95))
tau_post_CI
```
Now we'll find the posterior mean and posterior $90$ percent interval for $\psi$.

```{r}
psi_sim <- log((p1_sim/(1-p1_sim)) / (p2_sim/(1-p2_sim)))

# obtain posterior mean
psi_post_mean <- mean(psi_sim)
psi_post_mean

# obtain CI
psi_post_CI <- quantile(psi_sim, probs = c(0.05,0.95))
psi_post_CI
```