---
title: "13-11"
author: "Andrew Wu"
date: "2025-07-26"
output: html_document
---

Get the Coronary Risk-Factor Study (CORIS) data from the book website. Use backward stepwise logistic regression based on AIC to select a model. Summarize your results.

```{r}
url <- "https://www.stat.cmu.edu/~larry/all-of-statistics/=data/coris.dat"
destfile <- "coris.dat"
download.file(url,destfile)
```

```{r}
raw_data <- readLines(destfile)
head(raw_data)
```

We have to clean the file first.

```{r}
raw_data <- raw_data[-c(2:4)] # remove lines between header and data
```

Now we write the lines back into another file.

```{r}
writeLines(raw_data,"cleaned_coris.dat") # create a cleaned_coris.dat file to read data from
corisData <- read.table("cleaned_coris.dat",header=TRUE,sep=",")
head(corisData)
```

We need to run a backward stepwise logistic regression based on AIC to select a model, to predict CHD.

```{r}
# first we write a function that takes a dataframe and column indices for covariates, and returns the AIC of the logistic regression model
get_AIC <- function(df, response_idx, covariates) {
  
  # get a dataframe with just the columns we care about
  df_sub <- df[,c(response_idx,covariates)]
  
  head(df_sub)
  
  # fit a logistic model
  submodel <- glm(as.formula(paste(colnames(df_sub)[1], " ~ .")), data=df_sub, family=binomial)
  
  # get the AIC 
  R_AIC <- AIC(submodel)
  our_AIC <- R_AIC/(-2)
  
  return(our_AIC)
}
```

```{r}
# the function needs to start with all covariates, remove a covariate to get a larger AIC, and repeat until there's no way to remove another covariate without decreasing the AIC
# it will return a logistic model with these covariates
backward_stepwise_regression <- function(df,response_idx) {
  
  # total possible number of covariates; subtract one because one column is the response
  total_columns <- ncol(df)
  total_possible_covariates <- total_columns - 1
  
  # define a list of the covariate indices we're using
  current_covariates_indices <- c(1:total_columns)
  current_covariates_indices <- current_covariates_indices[current_covariates_indices != response_idx]
  
  # keep track of the AIC statistic value before removing a covariate
  current_AIC <- get_AIC(df, response_idx, current_covariates_indices)
  
  # for each possible removed covariate, we need to compute the AIC, and take the maximum among them
  # we'll set up a for loop
  for (i in 1:total_possible_covariates) {
    
    # we'll loop through current_covariates_indices
    # we'll compute the AIC upon removal of each element 
    max_AIC <- -Inf
    
    for (j in current_covariates_indices) {
      
      # compute the AIC with current_covariates_indices without j
      AIC <- get_AIC(df,response_idx,current_covariates_indices[current_covariates_indices != j]) 
      # we need to keep track of which column index leads to the maximum AIC
      max_AIC = max(AIC,max_AIC)
      if (max_AIC == AIC) {
        next_covariate <- j
      }
    }
    # we need to check whether max_AIC is actually greater than current
    if (max_AIC > current_AIC) {
      
      # if it is, then we need to reset current_AIC and remove from current_covariates_indices
      current_mallow <- max_AIC
      current_covariates_indices <- current_covariates_indices[current_covariates_indices != next_covariate]
    } 
    else {

      # if not, then we stop the backward regression
      predictors_col_names <- colnames(df)[current_covariates_indices]
      f <- as.formula(paste(colnames(df)[response_idx],"~",paste(predictors_col_names,collapse="+")))
      return(lm(f,data=df))
    }
  }
}
```

We need a dataframe without the first column, which just gives the row number.

```{r}
updatedCorisData <- corisData[-1]
```

Now we'll do the backwards stepwise regression.

```{r}
final_model <- backward_stepwise_regression(updatedCorisData,10)
summary(final_model)
```

So it seems the best predictors for coronary heart disease are tobacco usage, LDL, family history, Type A, and age. 




