---
title: "4-4"
author: "Andrew Wu"
date: "2025-02-21"
output: html_document
---

Let $X_1, \ldots , X_n \sim \text{Bernoulli}(p)$. Let $\alpha$ be fixed and define $\epsilon\_n = \sqrt{\frac{1}{2n} \log \left( \frac{2}{\alpha} \right)}$.

Let $\widehat{p}_n = n^{-1} \sum_{i=1}^n X_i$. Define $C_n = (\widehat{p}_n - \epsilon_n, \widehat{p}_n + \epsilon_n)$. We proved earlier with Hoeffding's inequality (in part (a)) that $\mathbb{P} (C_n \text{ contains } p) \ge 1 - \alpha$.

(b) Let's examine the properties of this confidence interval. Let $\alpha = 0.05$ and $p = 0.4$. Conduct a simulation study to see how often the interval contains $p$ (called the coverage.) Do this for various values of $n$ between $1$ and $10000$. Plot the coverage versus $n$.

(c) Plot the length of the interval versus $n$. Suppose we want the length of the interval to be no more than $.05$. How large should $n$ be?

```{r}
alpha = 0.05
p = 0.4
```

```{r}
# define a function that simulates runs confidence intervals based on n, alpha, p
simulate_CI <- function(n, alpha, p, runs) {
  
  # track how many times the CI contains the real p
  contains <- 0
  
  # compute eps_n before simulating; it depends only on n and alpha
  eps_n <- sqrt((1/(2*n)) * log(2/alpha))
  
  for (i in 1:runs) {
    
    # phat_n depends on the bernoulli variables we simulate
    phat_n <- mean(rbinom(n,1,p))

    # after generating phat_n, we can now compute the CIs
    CI <- c(phat_n - eps_n, phat_n + eps_n)
    
    # if p lies in the CI, increment contains
    if (p <= CI[2] && p >= CI[1]) {
      contains <- contains + 1
    }
  }
  return(contains/runs)
}
```

```{r}
# we'll do this for a few values of n
set.seed(126)

# the following lines print the proportion of generated CIs that p lies in
simulate_CI(10,alpha,p,1000)
simulate_CI(50,alpha,p,1000)
simulate_CI(250,alpha,p,1000)
simulate_CI(1000,alpha,p,1000)
simulate_CI(5000,alpha,p,1000)
simulate_CI(10000,alpha,p,1000)
```

Overall, the bound seems fairly conservative. The generated proportions are well greater than $1- \alpha = 0.95.$

```{r}
set.seed(126)
# to make sure the simulation doesn't take forever
# we'll just plot x = 1-9, 10-90, 100-900, 1000-9000
x_values <- c()
for (i in 0:3) {
  for (j in 1:9) {
    x_values <- c(x_values, j*10^i)
  }
}
x_values <- c(x_values, 10000)

# then create the y_values by applying the function with our established values of alpha and p
# runs is set to 1000; seems large enough
y_values <- sapply(x_values, function(x) simulate_CI(x,alpha,p,1000))
```

Here is a plot of the coverages versus $n$.

```{r}
plot(x_values,y_values, xlab="Value of n", ylab="Coverage", type="l")
```

Now, for part (c), we need to create a plot of interval length versus $n$. This depends only on $n$ and $\alpha$.

```{r}
simulate_interval_length <- function(n, alpha) {
  eps_n <- sqrt((1/(2*n)) * log(2/alpha))
  return(2*eps_n)
}
```

```{r}
x_values <- c(100:10000)
y_values <- sapply(x_values, function(x) simulate_interval_length(x, alpha))
```

```{r}
plot(x_values,y_values, xlab="Value of n", ylab="Length of confidence interval", type="l")
```
