---
title: "4-6"
author: "Andrew Wu"
date: "2025-02-21"
output: html_document
---

Let $Z \sim N(0,1)$. Plot $\mathbb{P}(|Z| > t)$ as a function of $t$. From Markov's inequality, we have the bound $\mathbb{P}(|Z| > t) \le \frac{\mathbb{E}(|Z|^k)}{t^k}$ for any $k > 0$. Plot these bounds for $k = 1, 2, 3, 4, 5$ and compare them to the true value of $\mathbb{P}(|Z| > t).$ Also, plot the bound from Mill's inequality.

Note that from previous work, we know that $\mathbb{P}(|Z| > t) = 2 \Phi(-t).$ We can use pnorm for the CDF of the normal distribution.

```{r}
x_values <- seq(0.1,5,by=0.1)
true_y_values <- sapply(x_values, function(x) 2*pnorm(-x))
```

```{r}
plot(x_values,true_y_values,xlab="t",ylab=expression("P(|Z| > t)"), type="l")
```

Now we'll plot the bounds from Markov's inequality.

```{r}
# the first five moments of |Z|
moments <- c(sqrt(2/pi), 1, 2*sqrt(2/pi), 3, 8*sqrt(2/pi))

# create the y-values by dividing moments by t^k
for (i in 1:5) {
  y_values <- sapply(x_values, function(x) moments[i]/(x^i))
  plot(x_values,y_values,xlab="t",ylab="Markov bound",type="l",ylim=c(0,10), col="blue")
  lines(x_values,true_y_values,col="red")
  legend("topright", legend=c("True Value","Markov bound"), col=c("red","blue"),lwd = 2, lty = 1)
}
```

Finally, we'll plot the bound from Mill's inequality.

```{r}
y_values <- sapply(x_values, function(x) sqrt(2/pi)*exp(-0.5*x^2)/x)
plot(x_values,y_values,xlab="t",ylab="Mill bound", type="l",ylim=c(0,5),col="blue")
lines(x_values,true_y_values, col="red")
legend("topright",legend=c("True Value","Mill bound"), col=c("red","blue"),lwd=2,lty=1)
```

