---
title: "7-3"
author: "Andrew Wu"
date: "2025-03-10"
output: html_document
---

Generate $100$ observations from an $N(0,1)$ distribution. Compute a $95$ percent confidence band for the CDF $F$. Repeat this $1000$ times and see how often the confidence band contains the true distribution function. Repeat using data from a Cauchy distribution.

We know from the textbook that a $1-\alpha$ Confidence Band for a CDF $F$ can be found by taking $L(x) = \max \{ \widehat{F}_n(x) - \epsilon_n, 0\}$, $U(x) = \min\{ \widehat{F}_n(x) + \epsilon_n, 1 \}$, with $\epsilon_n = \sqrt{\frac{1}{2n} \log (2/\alpha)}$. Then $\mathbb{P}(L(x) \le F(x) \le U(x) \text{ for all } x) \ge 1 - \alpha$.

```{r}
n <- 100 # need 100 observations
repeats <- 1000 # repeating 1000 times
alpha <- 0.05 # need the probability to be >= 95%, so need 1-alpha >= 0.95
eps_n <- sqrt(log(2/alpha)/(2*n)) # set epsilon_n as described previously
```

We want to check whether the CDF for $N(0,1)$, which we will from now on denote by $F$, lies between $L(x)$ and $U(x)$. We can do this for by checking whether $L(x) \le F(x) \le U(x)$ for all $100$ generated values of $x$; we repeat this procedure $1000$ times and see how many times it is true that $L(x) \le F(x) \le U(x)$ for all $100$ generated values.

```{r}
set.seed(127)

total_successes <- 0
F_nhat <- seq(1/100, 1, by = 1/100) # the estimated CDF for the generated values is just 1/100, 2/100, ... , 1 at those values

for (i in 1:repeats) {
  
  # set up the observations & confidence band
  
  generated_random_normals <- sort(rnorm(100)) # generate & sort 100 observations from N(0,1)
  F_values <- pnorm(generated_random_normals) # true N(0,1) CDF corresponding to generated values
  
  
  L_values <- pmax(F_nhat - eps_n,rep(0,n)) # pmax is vectorized & allows us to compare the generated - eps_n to a vector of 0s
  U_values <- pmin(F_nhat + eps_n,rep(1,n))
  
  
  # test whether confidence band contains the actual CDF
  
  success <- TRUE
  
  for (j in 1:n) {
    # check whether, for all the 
    if (F_values[j] < L_values[j] || F_values[j] > U_values[j]) {
      success <- FALSE
      break
    }
  }
  
  if (success) {
    total_successes <- total_successes + 1
  }
}

total_successes

```
So we find that in $969$ of the simulations, the true distribution lies within the $95\%$ confidence band. This makes sense; the DKW inequality states $\mathbb{P}(L(x) \le F(x) \le U(x) \text{ for all } x) \ge 1-\alpha$, so it's not unreasonable for us to be doing somewhat better than $1-\alpha$.

Now we just need to do the same thing for the Cauchy distribution.

```{r}
set.seed(127)

total_successes <- 0

for (i in 1:repeats) {
  
  # set up the observations & confidence band
  
  generated_random_cauchys <- sort(rcauchy(100)) # generate & sort 100 cauchy observations
  F_values <- pcauchy(generated_random_cauchys) # true cauchy CDF corresponding to generated values
  F_nhat <- seq(1/100, 1, by = 1/100) # the estimated CDF for the generated values is just 1/100, 2/100, ... , 1
  
  L_values <- pmax(F_nhat - eps_n,rep(0,n)) # pmax is vectorized & allows us to compare the generated - eps_n to a vector of 0s
  U_values <- pmin(F_nhat + eps_n,rep(1,n))
  
  
  # test whether confidence band contains the actual CDF
  
  success <- TRUE
  
  for (j in 1:n) {
    # check whether, for all the 
    if (F_values[j] < L_values[j] || F_values[j] > U_values[j]) {
      success <- FALSE
      break
    }
  }
  
  if (success) {
    total_successes <- total_successes + 1
  }
}

total_successes
```



