---
title: "7-8"
author: "Andrew Wu"
date: "2025-03-12"
output: html_document
---

Get the data on eruption times and waiting times between eruptions of the Old Faithful geyser from the website. Estimate the mean waiting time and give a standard error for the estimate. Also, give a $90$ percent confidence interval for the mean waiting time. Now estimate the median waiting time.

```{r}
url <- "https://stat.cmu.edu/~larry/all-of-statistics/=data/faithful.dat"
destfile <- "faithful.dat"
download.file(url,destfile)
```

```{r}
readLines(destfile,n=40) # checking for format
```

```{r}
faithfulData <- read.table("faithful.dat", header = TRUE, skip = 25) # skip = 25 helps us skip until the actual data
head(faithfulData)
```

A good estimate for the mean waiting time is just the average of all the waiting times.

```{r}
waitingTimes <- faithfulData$waiting # we don't actually work with the eruption times, so we just need the waiting times
X_nbar <- mean(waitingTimes)
X_nbar
```

We know that to estimate the standard error, we can divide the square root of the sample variance by the square root of the number of observations.

```{r}
n <- length(waitingTimes)
se_hat <- (sd(waitingTimes)) / (sqrt(n))
se_hat
```

A 90 percent confidence interval for the mean waiting time is given by $\overline{X}_n \pm z_{\alpha / 2}\widehat{\text{se}},$ with $\alpha = 0.1$.

```{r}
alpha <- 0.1
z <- qnorm(1 - alpha/2)
CI <- c(X_nbar - z*se_hat, X_nbar + z*se_hat)
CI
```

Finally, an estimate for the median waiting time is just given by $\widehat{F}_n^{-1}(0.5) = \inf \{x : \widehat{F}_n(x) \ge 0.5 \}$. In practice, that just means we can use median().

```{r}
median(waitingTimes)
```





