---
title: "8-2"
author: "Andrew Wu"
date: "2025-03-21"
output: html_document
---

Conduct a simulation to compare the various bootstrap confidence interval methods. Let $n = 50$ and let $T(F) = \int (x-\mu)^3 dF(x)/\sigma^3$ be the skewness. Draw $Y_1, \ldots , Y_n \sim N(0,1)$ and set $X_i = e^{Y_i}, i=1,\ldots, n$. Construct the three types of bootstrap $95$ percent intervals for $T(F)$ from the data $X_1, \ldots , X_n$. Repeat this whole thing many times and estimate the true coverage of the three intervals.

```{r}
n <- 50
Yis <- rnorm(n)
Xis <- exp(Yis)
```

We want to estimate the skewness $T(F)$ for the lognormal distribution, basically. We know from an example in the previous chapter that the plug-in estimator for the skewness is given by $\frac{\frac{1}{n} \sum_i (X_i - \widehat{\mu})^3}{\widehat{\sigma}^3}$.

```{r}
skew <- function(data) {
  n <- length(data)
  mu_data <- mean(data)
  sd_data <- sd(data)
  
  result <- sum(((data - mu_data) / sd_data)^3)/n 
  # R is vectorized, so this will go through all the elements of data and sum up the individual quantities
  
  return(result)
}
```

```{r}
skew_estimate <- skew(Xis)
```

Now we bootstrap.

```{r}
B <- 1000
skews <- c()

for (i in 1:B) {
  sampled_data <- sample(Xis,n,replace=TRUE)
  skews[i] <- skew(sampled_data)
}
```

Now the vector skews has all our bootstrapped skews. We can construct our confidence intervals.

For the normal, we also need the bootstrapped standard error.

```{r}
se_boot <- sd(skews)
alpha <- 0.05
se_boot
normal_ci <- c(skew_estimate - qnorm(1-alpha/2)*se_boot, skew_estimate + qnorm(1-alpha/2)*se_boot)
normal_ci
```

Next, the pivotal:

```{r}
pivotal_ci <- c(2*skew_estimate - quantile(skews,1-alpha/2)[[1]], 2*skew_estimate - quantile(skews,alpha/2)[[1]])
pivotal_ci
```

Finally, the percentile:

```{r}
percentile_ci <- c(quantile(skews,alpha/2)[[1]], quantile(skews,1-alpha/2)[[1]])
percentile_ci
```

Now we must repeat all of this many times and see how often each interval contains the true skewness, which we know to be $(e+2) \sqrt{e - 1}$.

```{r}
true_skewness <- (exp(1) + 2)*sqrt(exp(1) - 1)
skews <- c()
normal_empirical_coverages <- c()
pivotal_empirical_coverages <- c()
percentile_empirical_coverages <- c()
```

```{r}
repetitions <- 1000
for (i in 1:repetitions) {
  
  # begin by drawing the Yis and setting the Xis
  Yis <- rnorm(n)
  Xis <- exp(Yis)
  
  # estimate the skew with the drawn and set data
  skew_estimate <- skew(Xis)
  
  # bootstrap
  # note we previously set B=1000 and n=50
  for (j in 1:B) {
    sampled_data <- sample(Xis,n,replace=TRUE)
    skews[j] <- skew(sampled_data)
  }
  
  # get the standard error
  se_boot <- sd(skews)
  
  # get the confidence intervals
  normal_ci <- c(skew_estimate - qnorm(1-alpha/2)*se_boot, skew_estimate + qnorm(1-alpha/2)*se_boot)
  pivotal_ci <- c(2*skew_estimate - quantile(skews,1-alpha/2)[[1]], 2*skew_estimate - quantile(skews,alpha/2)[[1]])
  percentile_ci <- c(quantile(skews,alpha/2)[[1]], quantile(skews,1-alpha/2)[[1]])
  
  # check whether the true value is in the CIs
  normal_empirical_coverages[i] <- (true_skewness >= normal_ci[1]) & (true_skewness <= normal_ci[2])
  pivotal_empirical_coverages[i] <- (true_skewness >= pivotal_ci[1]) & (true_skewness <= pivotal_ci[2])
  percentile_empirical_coverages[i] <- (true_skewness >= percentile_ci[1]) & (true_skewness <= percentile_ci[2])
}
```

```{r}
mean(normal_empirical_coverages)
mean(pivotal_empirical_coverages)
mean(percentile_empirical_coverages)
```













