---
title: "8-3"
author: "Andrew Wu"
date: "2025-03-23"
output: html_document
---

Let $X_1,\ldots , X_n \sim t_3$, where $n = 25$. Let $\theta = T(F) = (q_{.75} - q_{.25})/1.34$ where $q_p$ denotes the $p$th quantile. Do a simulation to compare the coverage and length of the following confidence intervals for $\theta$: (i) Normal interval with standard error from the bootstrap, (ii) bootstrap percentile interval, and (iii) pivotal bootstrap interval.

```{r}
set.seed(127)

n <- 25 # number of observations
df <- 3 # degrees of freedom

Xis <- rt(n,df)
```

Now we compute an estimate of for this data.

```{r}
q <- quantile(Xis,c(0.25,0.75)) # obtain q_{.25} and q_{.75}
theta_hat <- (q[[2]] - q[[1]])/1.34
theta_hat
```

To get the confidence intervals, we first bootstrap.

```{r}
B <- 1000 
theta_hats <- c()

for (i in 1:B) {
  sampled_data <- sample(Xis,n,replace=TRUE) # draw n samples from Xis with replacement
  q <- quantile(sampled_data,c(0.25,0.75)) # compute the 0.25 and 0.75 quantiles
  theta_hats[i] <- (q[[2]] - q[[1]])/1.34 # compute the statistic for the sample
}
```

The standard error is the standard deviation of the thetas.

```{r}
se_boot <- sd(theta_hats)
se_boot
```

For the normal confidence interval:

```{r}
alpha <- 0.05
normal_ci <- c(theta_hat - qnorm(1-alpha/2)*se_boot, theta_hat + qnorm(1-alpha/2)*se_boot)
normal_ci
```

The percentile:

```{r}
percentile_ci <- c(quantile(theta_hats,alpha/2)[[1]], quantile(theta_hats,1-alpha/2)[[1]])
percentile_ci
```

The pivotal:

```{r}
pivotal_ci <- c(2*theta_hat - quantile(theta_hats,1-alpha/2)[[1]], 2*theta_hat - quantile(theta_hats,alpha/2)[[1]])
pivotal_ci
```

Now note that our task is to run a simulation to see how often these intervals contain the true value. Accordingly, we first find the true value of the statistic, and then simulate.

```{r}
theta <- (qt(0.75,df) - qt(0.25,df))/1.34
theta
```

```{r}
repetitions <- 5000

# track coverages
normal_empirical_coverages <- c()
pivotal_empirical_coverages <- c()
percentile_empirical_coverages <- c()

for (i in 1:repetitions) {
  
  # draw 25 observations from t_3. note we set n and df earlier
  Xis <- rt(n,df)
  
  # get theta_hat
  q <- quantile(Xis,c(0.25,0.75))
  theta_hat <- (q[[2]]-q[[1]])/1.34
  
  # bootstrap
  theta_hats <- c()
  for (j in 1:B) {
    sampled_data <- sample(Xis, n, replace=TRUE)
    q <- quantile(sampled_data,c(0.25,0.75))
    theta_hats[j] <- (q[[2]] - q[[1]])/1.34
  }
  
  # get the standard error
  se_boot <- sd(theta_hats)
  
  # create confidence intervals
  normal_ci <- c(theta_hat - qnorm(1-alpha/2)*se_boot, theta_hat + qnorm(1-alpha/2)*se_boot)
  percentile_ci <- c(quantile(theta_hats,alpha/2)[[1]], quantile(theta_hats,1-alpha/2)[[1]])
  pivotal_ci <- c(2*theta_hat - quantile(theta_hats,1-alpha/2)[[1]], 2*theta_hat - quantile(theta_hats,alpha/2)[[1]])

  # check whether the true value is in the CIs
  normal_empirical_coverages[i] <- (theta >= normal_ci[1]) & (theta <= normal_ci[2])
  pivotal_empirical_coverages[i] <- (theta >= pivotal_ci[1]) & (theta <= pivotal_ci[2])
  percentile_empirical_coverages[i] <- (theta >= percentile_ci[1]) & (theta <= percentile_ci[2])
}
```

```{r}
mean(normal_empirical_coverages)
mean(pivotal_empirical_coverages)
mean(percentile_empirical_coverages)
```













