---
title: "8-6"
author: "Andrew Wu"
date: "2025-03-23"
output: html_document
---

Let $X_1, \ldots , X_n \sim \text{Normal}(\mu, 1)$. Let $\theta = e^\mu$ and let $\widehat{\theta} = e^{\overline{X}}$. Create a data set, using $\mu = 5$, consisting of $n = 100$ observations.

(a) Use the bootstrap to get the se and $95$ percent confidence interval for $\theta$.

(b) Plot a histogram of the bootstrap replications. This is an estimate of the distribution of $\widehat{\theta}$. Compare this to the true sampling distribution of $\widehat{\theta}.$

```{r}
set.seed(127)

n <- 100
mu <- 5
Xis <- rnorm(n,mu,1)
theta <- exp(mu)
mu_hat <- mean(Xis)
theta_hat <- exp(mu_hat)
```

Now we bootstrap.

```{r}
B <- 100
sampled_theta_hats <- c()

for (i in 1:B) {
  sampled_data <- sample(Xis,n,replace=TRUE) # draw n times from Xis with replacement
  sampled_theta_hats[i] <- exp(mean(sampled_data))
}

se_hat <- sd(sampled_theta_hats)
se_hat
```

The problem doesn't really specify what type of $95$ percent confidence interval we're looking for, so I will give the percentile interval.

```{r}
alpha <- 0.05
ci_percentile <- c(quantile(sampled_theta_hats,alpha/2)[[1]],quantile(sampled_theta_hats,1-alpha/2)[[1]])
ci_percentile
```

Here is a histogram of the bootstrap replications.

```{r}
hist(sampled_theta_hats)
```

We can find the true distribution of $\widehat{\theta}$ by noting that (1) $\overline{X}_n \sim N(\mu, \frac{1}{n})$, and (2) $\widehat{\theta} \sim \exp (\overline{X}_n)$. It follows that $\widehat{\theta}$ follows a lognormal distribution.

```{r}
curve(dlnorm(x, meanlog=mu, sdlog = sqrt(1/n)),
      from = 100, to = 200
      )
```

The distributions look reasonably similar.