---
title: "8-7"
author: "Andrew Wu"
date: "2025-03-23"
output: html_document
---

Let $X_1, \ldots , X_n \sim \text{Uniform}(0,\theta)$. Let $\widehat{\theta} = \max \{X_1, \ldots , X_n\}$. Generate a data set of size $50$ with $\theta = 1$. Compare the true distribution of $\widehat{\theta}$ to the histogram from the bootstrap.

```{r}
set.seed(127)
n <- 50
theta <- 1
Xis <- runif(n,0,theta)
theta_hat <- max(Xis)
theta_hat
```

We find (elsewhere) that the true distribution of $\widehat{\theta}$ is $f(x) = 50x^{49}$ for $0 \le x \le 1$. Therefore, we'll bootstrap and then compare the histogram to this distribution.

```{r}
B <- 1000
theta_hats <- c()

for (i in 1:B) {
  sampled_data <- sample(Xis, n, replace=TRUE)
  theta_hats[i] <- max(sampled_data)
}

hist(theta_hats,
     breaks=20,
     main="Histogram of bootstrapped estimates for theta")
```

```{r}
curve(50*x^{49}, from = 0.85, to = 1,
      main = "True distribution of theta",
      ylab = "f(x)")
```

Here the true distribution seems a lot thicker near $1$, compared to the bootstrapped histogram.
