---
title: "9-10"
author: "Andrew Wu"
date: "2025-04-13"
output: html_document
---

Let $X_1, \ldots , X_n \sim \text{Uniform}(0,\theta)$. The MLE is $\widehat{\theta} = \max \{X_1, \ldots , X_n\}$. Generate a dataset of size $50$ with $\theta = 1$.

Compare the true distribution of $\widehat{\theta}$ to the histogram from the parametric bootstrap.

Show that for the parametric bootstrap $\mathbb{P}(\widehat{\theta}^* = \widehat{\theta} = 0)$.

(Note that most of this problem was completed previously in chapter 8.)

```{r}
set.seed(127)
n <- 50
theta <- 1
Xis <- runif(n,0,theta)
theta_hat <- max(Xis)
theta_hat
```

Now we create the parametric bootstrap.

```{r}
B <- 10000
thetas_parametric <- numeric(B)

for (i in 1:B) {
  Xis_parametric <- runif(n,0,theta_hat) # use the MLE of theta
  thetas_parametric[i] <- max(Xis_parametric)
}

hist(thetas_parametric)
```

```{r}
curve(50*x^{49}, from = 0.85, to = 1,
      main = "True distribution of theta",
      ylab = "f(x)")
```

These distributions are quite similar.

Regarding the other part of the problem, we have $\mathbb{P}(\widehat{\theta}^* = \widehat{\theta} = 0)$ because $\widehat{\theta}^*$ is the maximum among samples drawn from $\text{Uniform}(0,\widehat{\theta})$ and $\mathbb{P}(X = \widehat{\theta}) = 0$ when $X \sim \text{Uniform}(0,\widehat{\theta})$.
