---
title: "9-7"
author: "Andrew Wu"
date: "2025-04-12"
output: html_document
---

$n_1$ people are given treatment $1$ and $n_2$ people are given treatment $2$. Let $X_1$ be the number of people on treatment $1$ who respond favorably to the treatment and let $X_2$ be the number of people on treatment $2$ who respond favorably. Assume that $X_1 \sim \bin(n_1,p_1)$ and $X_2 \sim \bin(n_2,p_2)$. Let $\psi = p_1 - p_2$ with MLE $\widehat{\psi}$.

Suppose $n_1 = n_2 = 200, X_1 = 160,$ and $X_2 = 148$. Find $\widehat{\psi}$. Find an approximate $90$ percent confidence interval for $\psi$ using the delta method and using the parametric bootstrap.

```{r}
n1 <- 200
n2 <- 200
X1 <- 160
X2 <- 148

# we already computed a general form for psi_hat in the other document
p1_hat <- X1/n1
p2_hat <- X2/n2
psi_hat <- p1_hat - p2_hat

psi_hat
```

To find a $90$ percent confidence interval using the delta method, we'll use the standard error we just computed.

```{r}
se_hat <- sqrt(p1_hat*(1 - p1_hat) / n1 + p2_hat*(1 - p2_hat) / n2)
CI <- c(psi_hat - qnorm(0.95)*se_hat, psi_hat + qnorm(0.95)*se_hat)
CI
```

Now we'll find the $90$ percent confidence interval using the bootstrap.

```{r}
rbinom(7,n1,p1_hat)
```

```{r}
B <- 10000

# don't actually need a for loop this time with rbinom
bootstrap_X1s <- rbinom(B,n1,p1_hat)
bootstrap_X2s <- rbinom(B,n2,p2_hat)

# use psi = p1 - p2
bootstrap_psis <- (bootstrap_X1s - bootstrap_X2s)/n1 # need probabilities, so can't just simply subtract
bootstrap_se <- sd(bootstrap_psis)
bootstrap_se
```

Now we compute the confidence interval similarly.

```{r}
CI_bootstrap <- c(psi_hat - qnorm(0.95)*bootstrap_se, psi_hat + qnorm(0.95)*bootstrap_se)
CI_bootstrap
```



