---
title: "9-9"
author: "Andrew Wu"
date: "2025-04-13"
output: html_document
---

Let $X_1, \ldots , X_n \sim N(\mu, 1)$. Let $\theta = e^\mu$ and let $\widehat{\theta} = e^{\overline{X}}$ be the MLE. Using $\mu = 5$, create a data set consisting of $n = 100$ observations.

(a) Use (i) the delta method, (ii) the parametric bootstrap, and (iii) the nonparametric bootstrap to get $\widehat{\text{se}}$ and a $95$ percent confidence interval for $\theta$.

(b) Plot a histogram of the bootstrap replications for the parametric and nonparametric bootstraps. These are estimates of the distribution of $\widehat{\theta}$. The delta method also gives an approximation to this distribution: namely, $N(\widehat{\theta}, \text{se}^2)$. Compare these to the true sampling distribution of $\widehat{\theta}$, which you can get by simulation. Which approximation is closer to the true distribution?

For part (a): We computed separately that the delta method gives $\widehat{\text{se}}(\widehat{\theta}) = e^{\overline{X}} \cdot \sqrt{1/n}$.

```{r}
set.seed(127)
n <- 100
mu <- 5

# compute SE with delta method
Xis <- rnorm(n,mu,1)
X_bar <- mean(Xis)
se_delta <- exp(X_bar) / sqrt(n)
se_delta

# get a 95% confidence interval
theta_hat <- exp(X_bar)
CI_delta <- c(theta_hat - qnorm(0.975)*se_delta, theta_hat + qnorm(0.975)*se_delta)
CI_delta
```

Next, we'll use the parametric bootstrap.

```{r}
set.seed(127)

B <- 10000
thetas_parametric <- numeric(B)

for (i in 1:B) {
  Xis_bootstrap <- rnorm(n,X_bar,1) # draw n samples from N(X_bar, 1)
  theta_bootstrap <- exp(mean(Xis_bootstrap)) # compute the bootstrapped estimator for theta
  thetas_parametric[i] <- theta_bootstrap # store in thetas_parametric
}

se_parametric <- sd(thetas_parametric)
se_parametric

CI_parametric_normal <- c(theta_hat - qnorm(0.975)*se_parametric, theta_hat + qnorm(0.975)*se_parametric)
CI_parametric_normal

CI_parametric_percentile <- quantile(thetas_parametric,c(0.025,0.975))
CI_parametric_percentile
```

And finally, the nonparametric bootstrap.

```{r}
set.seed(127)
thetas_nonparametric <- numeric(B)

for (i in 1:B) {
  Xis_bootstrap <- sample(Xis,n,replace=TRUE) # draw n samples from the Xis with replacement
  theta_bootstrap <- exp(mean(Xis_bootstrap)) # compute the bootstrapped estimator for theta
  thetas_nonparametric[i] <- theta_bootstrap # store in thetas_nonparametric
}

se_nonparametric <- sd(thetas_nonparametric)
se_nonparametric

CI_nonparametric_normal <- c(theta_hat - qnorm(0.975)*se_nonparametric, theta_hat + qnorm(0.975)*se_nonparametric)
CI_nonparametric_normal

CI_nonparametric_percentile <- quantile(thetas_nonparametric, c(0.025,0.975))
CI_nonparametric_percentile
```

For part (b), we will plot the histograms.

```{r}
hist(thetas_parametric, xlim=c(100,240), breaks=14)
hist(thetas_nonparametric, xlim=c(100,240), breaks=14)
```

Then, the delta method approximation.

```{r}
curve(dnorm(x, theta_hat, se_delta), from = 100, to = 240)
```

Finally, we'll simulate the distribution of $\widehat{\theta}$.

```{r}
set.seed(127)
thetas_simulated <- numeric(B)

for (i in 1:B) {
  Xis_simulated <- rnorm(n, mu, 1)
  X_bar_simulated <- mean(Xis_simulated)
  thetas_simulated[i] <- exp(X_bar_simulated)
}

hist(thetas_simulated, xlim=c(100,240), breaks=14)
```

They're all quite similar. 